---
title: "3. Collecting Gunner Data"
output:
  html_document:
    df_print: paged
---

In this notebook, I will be collecting the necessary data for my models. This includes pulling out existing variables from the data and deriving new features.

## 3.1 Imports
```{r}
source("E:/R Projects/gunner_evaluation/R/00_source.R")


library(tidyverse)

puntPlays <- read.csv(paste0(wd$data, 'punt_plays1.csv'))
puntPlayInfo <- read.csv(paste0(wd$data, 'punt_play_info.csv'))
```

## 3.2 Create New Dataframe to Store Gunner Data
This dataframe will store the new features derived from the tracking data and existing features.

```{r}
specialistData <- data.frame(gameId = rep(puntPlayInfo$gameId, each = 2),
                             playId = rep(puntPlayInfo$playId, each = 2),
                             gunnerTeam = character(2 * nrow(puntPlayInfo)),
                             gunnerId = numeric(2 * nrow(puntPlayInfo)),
                             gunnerName = character(2 * nrow(puntPlayInfo)),
                             viseTeam = character(2 * nrow(puntPlayInfo)),
                             viseId = numeric(2 * nrow(puntPlayInfo)),
                             viseName = character(2 * nrow(puntPlayInfo)),
                             firstManDown = numeric(2 * nrow(puntPlayInfo)),
                             timeToBeatVise = numeric(2 * nrow(puntPlayInfo)),
                             disFromLOS = numeric(2 * nrow(puntPlayInfo)),
                             disFromReturner = numeric(2 * nrow(puntPlayInfo)),
                             speedDev = numeric(2 * nrow(puntPlayInfo)),
                             topSpeed = numeric(2 * nrow(puntPlayInfo)),
                             tackle = numeric(2 * nrow(puntPlayInfo)),
                             missedTackle = numeric(2 * nrow(puntPlayInfo)),
                             squeezeDis = numeric(2 * nrow(puntPlayInfo)),
                             gunnerSide = numeric(2 * nrow(puntPlayInfo)),
                             release = numeric(2 * nrow(puntPlayInfo)),
                             kickDirRelGun = numeric(2 * nrow(puntPlayInfo)),
                             cat = numeric(2 * nrow(puntPlayInfo)),
                             correctRelease = numeric(2 * nrow(puntPlayInfo)))
```

Join with punt play info
```{r}
specialistData <- specialistData %>%
  inner_join(puntPlayInfo)
```

## 3.3 Store Identifying Information for Gunners and Jammers (Team, Name, Jersey Number, ID)
```{r}
for (i in seq(1, nrow(specialistData), 2)) {
  play <- puntPlays %>% filter(gameId == specialistData$gameId[i] & 
                                 playId == specialistData$playId[i])
  gunners <- play$gunner[1]
  gunnerTeam <- play$possessionTeam[1]
  gunner1 <- unlist(strsplit(gunners, split = '; '))[1]
  gunner2 <- unlist(strsplit(gunners, split = '; '))[2]
  gunner1Number <- substrRight(gunner1, 2)
  gunner2Number <- substrRight(gunner2, 2)
  
  specialistData$gunnerId[i] <- play$nflId[which(play$teamAbbr == gunnerTeam & 
                                                   play$jerseyNumber == gunner1Number)][1]
  specialistData$gunnerId[i + 1] <- play$nflId[which(play$teamAbbr == gunnerTeam & 
                                                       play$jerseyNumber == gunner2Number)][1]
  specialistData$gunnerName[i] <- play$displayName[which(play$teamAbbr == gunnerTeam & 
                                                           play$jerseyNumber == gunner1Number)][1]
  specialistData$gunnerName[i + 1] <- play$displayName[which(play$teamAbbr == gunnerTeam & 
                                                               play$jerseyNumber == gunner2Number)][1]
  specialistData$gunnerTeam[c(i, i + 1)] <- gunnerTeam
}
```

```{r}
for (i in seq(1, nrow(specialistData), 2)) {
  play <- puntPlays %>% filter(gameId == specialistData$gameId[i] & 
                                 playId == specialistData$playId[i])
  vises <- play$vises[1]
  viseTeam <- gsub(" .*$", "", vises)
  vise1 <- unlist(strsplit(vises, split = '; '))[1]
  vise2 <- unlist(strsplit(vises, split = '; '))[2]
  vise1Number <- substrRight(vise1, 2)
  vise2Number <- substrRight(vise2, 2)
  
  specialistData$viseId[i] <- play$nflId[which(play$teamAbbr == viseTeam & 
                                                 play$jerseyNumber == vise1Number)][1]
  specialistData$viseId[i + 1] <- play$nflId[which(play$teamAbbr == viseTeam & 
                                                     play$jerseyNumber == vise2Number)][1]
  specialistData$viseName[i] <- play$displayName[which(play$teamAbbr == viseTeam & 
                                                         play$jerseyNumber == vise1Number)][1]
  specialistData$viseName[i + 1] <- play$displayName[which(play$teamAbbr == viseTeam & 
                                                             play$jerseyNumber == vise2Number)][1]
  specialistData$viseTeam[c(i, i + 1)] <- viseTeam
}
```

```{r}
head(specialistData)
```

### 3.3.1 Remove Plays with Missing Player IDs
There were some instances where a gunner would be listed, say 'CLE 27', in the PFF data, but there would be no #27 for Cleveland in the tracking data. These plays were removed.
```{r}
missingIDs <- which(is.na(specialistData$gunnerId) | is.na(specialistData$viseId))
gameID <- specialistData$gameId
playID <- specialistData$playId
rowsRemove <- numeric(0)
for (i in missingIDs) {
  if (gameID[i] == gameID[i + 1] & playID[i] == playID[i + 1]) {
    rowsRemove <- c(rowsRemove, i,  i + 1)
  } else {
    rowsRemove <- c(rowsRemove, i, i - 1)
  }
}
specialistData <- specialistData[-rowsRemove, ]

sum(is.na(specialistData$gunnerId))
```

## 3.4 Remove Odd Plays
There were 3 odd plays that I'd noticed in the data. One play had 2 jammers fielded, but they were both lined up on the same side of the field. Another play listed 2 jammers, but clearly had 3 jammers fielded. Finally, the other play's tracking data began after the ball was snapped. These plays were removed.
```{r}
specialistData <- specialistData[-c(2049:2050, 2673:2674, 3943:3944), ]
```

## 3.5 Match Gunners to Jammer Guarding them
I have each gunner and jammer from each play, but they might not be lined up correctly. I want the dataframe to be set up such that the jammer in row n is defending the gunner in row n during that play. This is possible by finding the distance between a gunner to each jammer, and moving them to the same row as the jammer they are closer to. This will make calculations in the future easier.
```{r}
for (i in seq(1, nrow(specialistData), 2)) {
  gunnerID <- specialistData$gunnerId[i]
  vise1ID <- specialistData$viseId[i]
  vise2ID <- specialistData$viseId[i + 1]
  vise1Name <- specialistData$viseName[i]
  vise2Name <- specialistData$viseName[i + 1]
  
  playStartRows <- which(puntPlays$gameId == specialistData$gameId[i] & 
                           puntPlays$playId == specialistData$playId[i] & 
                           puntPlays$frameId == 1)
  
  # first frame of play
  playStart <- puntPlays[playStartRows, ]
  
  # distance between gunner and jammer 1
  dis1 <- abs(as.numeric(playStart[which(playStart$nflId == gunnerID), "y"]) - 
                as.numeric(playStart[which(playStart$nflId == vise1ID), "y"]))
  
  # distance between gunner and jammer 2
  dis2 <- abs(as.numeric(playStart[which(playStart$nflId == gunnerID), "y"]) - 
                as.numeric(playStart[which(playStart$nflId == vise2ID), "y"]))
  
  # compare distance
  if (dis2 < dis1) {
    specialistData$viseId[i] <- vise2ID
    specialistData$viseName[i] <- vise2Name
    specialistData$viseId[i + 1] <- vise1ID
    specialistData$viseName[i + 1] <- vise1Name
  }
}

head(specialistData)
```

## 3.6 Fill *returnYds* and *specialTeamsResult* Columns
```{r}
returnYds <- numeric(nrow(specialistData))
for (i in seq(1, nrow(specialistData), 2)) {
  play <- puntPlays %>% filter(gameId == specialistData$gameId[i] &
                               playId == specialistData$playId[i])
  
  specialistData$specialTeamsResult[c(i, i + 1)] <- play$specialTeamsResult[1]
  returnYds[c(i, i + 1)] <- play$kickReturnYardage[1]
}
specialistData$returnYds <- returnYds

head(specialistData)
```

## 3.7 Fill *timeToBeatVise* Column
This is the time it takes the gunner to get past the jammer defending them in seconds. The gunner must then be able to continuously stay past the jammer up until the ball is caught. If a gunner is able to get past their jammer quickly, they have a straight shot at the returner. This should either cause a fair catch or allow the gunner to make a quick tackle.
```{r}
for (i in seq(1, nrow(specialistData), 2)) {
  play <- puntPlays %>% filter(gameId == specialistData$gameId[i] &
                               playId == specialistData$playId[i])
  
  # find rows containing each player
  returnerRows <- which(play$nflId == play$returnerId)
  gunner1Rows <- which(play$nflId == specialistData$gunnerId[i])
  vise1Rows <- which(play$nflId == specialistData$viseId[i])
  gunner2Rows <- which(play$nflId == specialistData$gunnerId[i + 1])
  vise2Rows <- which(play$nflId == specialistData$viseId[i + 1])
  
  # for each frame, store gunner's distance to returner and jammer's distance to returner
  framesGunner1PastVise1 <- numeric(0)
  for (j in 1:length(gunner1Rows)) {
    gunner1DisReturner <- eucDis(play[gunner1Rows[j], c("x", "y")], play[returnerRows[j], c("x", "y")])
    vise1DisReturner <- eucDis(play[vise1Rows[j], c("x", "y")], play[returnerRows[j], c("x", "y")])
    
    # store frames where gunner is closer to returner (past jammer)
    if (gunner1DisReturner < vise1DisReturner) {
      framesGunner1PastVise1 <- c(framesGunner1PastVise1, j)
    }
  }
  
  # split frames where gunner is past jammer into sequences and save last sequence
  sequentialFrames1 <- split(framesGunner1PastVise1, cumsum(c(TRUE, diff(framesGunner1PastVise1) != 1)))
  specialistData$timeToBeatVise[i] <- sequentialFrames1[[length(sequentialFrames1)]][1] / 10
  
  # repeat for other gunner
  framesGunner2PastVise2 <- numeric(0)
  for (j in 1:length(gunner2Rows)) {
    gunner2DisReturner <- eucDis(play[gunner2Rows[j], c("x", "y")], play[returnerRows[j], c("x", "y")])
    vise2DisReturner <- eucDis(play[vise2Rows[j], c("x", "y")], play[returnerRows[j], c("x", "y")])
    
    if (gunner2DisReturner < vise2DisReturner) {
      framesGunner2PastVise2 <- c(framesGunner2PastVise2, j)
    }
  }
  
  sequentialFrames2 <- split(framesGunner2PastVise2, cumsum(c(TRUE, diff(framesGunner2PastVise2) != 1)))
  specialistData$timeToBeatVise[i + 1] <- sequentialFrames2[[length(sequentialFrames2)]][1] / 10
}

head(specialistData)
```

## 3.8 Fill *firstManDown* Column
This variable identifies which gunner was the first man down the field.
```{r}
for (i in seq(1, nrow(specialistData), 2)) {
  comp <- c(specialistData$timeToBeatVise[i], specialistData$timeToBeatVise[i + 1])
  
  if (sum(is.na(comp)) == 0) {
    specialistData$firstManDown[i + (which(comp == min(comp)) - 1)] <- 1
  } else if (sum(is.na(comp)) == 1) {
    specialistData$firstManDown[i + (which(!is.na(comp)) - 1)] <- 1
  }
}
```


## 3.9 Fill *disFromLOS* Column
This is the gunner's distance from the line of scrimmage at the point of the ball being caught. A gunner who is able to make it further down the field should either be able to force a fair catch, or make a tackle quicker.
```{r}
for (i in seq(1, nrow(specialistData), 2)) {
  play <- puntPlays %>% filter(gameId == specialistData$gameId[i] &
                               playId == specialistData$playId[i])
  
  LOS <- play$absoluteYardlineNumber[1]
  gunner1 <- specialistData$gunnerId[i]
  gunner2 <- specialistData$gunnerId[i + 1]
  
  snapFrame <- puntPlayInfo$snapFrame[which(puntPlayInfo$gameId == play$gameId[1] &
                                              puntPlayInfo$playId == play$playId[1])]
  
  catchFrame <- puntPlayInfo$catchFrame[which(puntPlayInfo$gameId == play$gameId[1] &
                                                puntPlayInfo$playId == play$playId[1])]
  
  # Row indices when ball is caught
  caughtRow <- play[which(play$frameId == catchFrame), c('nflId', 'x')]
  if (nrow(caughtRow) == 0) {
    caughtRow <- play[which(play$frameId == max(play$frameId)), c('nflId', 'x')]
  }

  # player pos when ball is caught
  gunner1Pos <- caughtRow[which(caughtRow$nflId == gunner1), 2]
  gunner2Pos <- caughtRow[which(caughtRow$nflId == gunner2), 2]
  
  specialistData$disFromLOS[i] <- abs(gunner1Pos - LOS)
  specialistData$disFromLOS[i + 1] <- abs(gunner2Pos - LOS)
}
```


## 3.10 Fill *disFromReturner* Column
This is the gunner's distance from the returner at the point of the ball being caught.
```{r}
for (i in seq(1, nrow(specialistData), 2)) {
  play <- puntPlays %>% filter(gameId == specialistData$gameId[i] &
                               playId == specialistData$playId[i])
  
  # ID players
  returner <- play$returnerId[1]
  gunner1 <- specialistData$gunnerId[i]
  gunner2 <- specialistData$gunnerId[i + 1]
  
  snapFrame <- puntPlayInfo$snapFrame[which(puntPlayInfo$gameId == play$gameId[1] &
                                              puntPlayInfo$playId == play$playId[1])]
  
  catchFrame <- puntPlayInfo$catchFrame[which(puntPlayInfo$gameId == play$gameId[1] &
                                                puntPlayInfo$playId == play$playId[1])]
  
  # Row indices when ball is caught
  caughtRow <- play[which(play$frameId == catchFrame), c('nflId', 'x', 'y')]
  if (nrow(caughtRow) == 0) {
    caughtRow <- play[which(play$frameId == max(play$frameId)), c('nflId', 'x', 'y')]
  }
  
  # player pos when ball is caught
  gunner1Pos <- caughtRow[which(caughtRow$nflId == gunner1), c("x", "y")]
  gunner2Pos <- caughtRow[which(caughtRow$nflId == gunner2), c("x", "y")]
  returnerPos <- caughtRow[which(caughtRow$nflId == returner), c("x", "y")]
  
  specialistData$disFromReturner[i] <- as.numeric(eucDis(gunner1Pos, returnerPos))
  specialistData$disFromReturner[i + 1] <- as.numeric(eucDis(gunner2Pos, returnerPos))
}
```


## 3.11 Fill *speedDev* Column
This is the standard deviation of a gunner's speed up until the point of the ball being caught. My thinking is that a more consistent speed indicates that the gunner is skilled or strong enough to keep their top speed while the jammer pushes and pulls on them.

```{r}
for (i in seq(1, nrow(specialistData), 2)) {
  play <- puntPlays %>% filter(gameId == specialistData$gameId[i] &
                               playId == specialistData$playId[i])
  
  gunner1 <- specialistData$gunnerId[i]
  gunner2 <- specialistData$gunnerId[i + 1]
  
  snapFrame <- puntPlayInfo$snapFrame[which(puntPlayInfo$gameId == play$gameId[1] &
                                              puntPlayInfo$playId == play$playId[1])]
  
  catchFrame <- puntPlayInfo$catchFrame[which(puntPlayInfo$gameId == play$gameId[1] &
                                                puntPlayInfo$playId == play$playId[1])]
  
  # gunner tracking data after ball is snapped and before ball is caught
  gunner1Frames <- play %>% filter(frameId >= snapFrame & frameId <= catchFrame & nflId == gunner1)
  gunner2Frames <- play %>% filter(frameId >= snapFrame & frameId <= catchFrame & nflId == gunner2)
  
  specialistData$speedDev[i] <- sd(gunner1Frames$s)
  specialistData$speedDev[i + 1] <- sd(gunner2Frames$s)
}
```


## 3.12 Fill *topSpeed* Column
This tracks a gunner's top speed up until the ball is caught.
```{r}
for (i in seq(1, nrow(specialistData), 2)) {
  play <- puntPlays %>% filter(gameId == specialistData$gameId[i] &
                               playId == specialistData$playId[i])
  
  gunner1 <- specialistData$gunnerId[i]
  gunner2 <- specialistData$gunnerId[i + 1]
  
  snapFrame <- puntPlayInfo$snapFrame[which(puntPlayInfo$gameId == play$gameId[1] &
                                              puntPlayInfo$playId == play$playId[1])]
  
  catchFrame <- puntPlayInfo$catchFrame[which(puntPlayInfo$gameId == play$gameId[1] &
                                                puntPlayInfo$playId == play$playId[1])]
  
  # gunner tracking data after ball is snapped and before ball is caught
  gunner1Frames <- play %>% filter(frameId >= snapFrame & frameId <= catchFrame & nflId == gunner1)
  gunner2Frames <- play %>% filter(frameId >= snapFrame & frameId <= catchFrame & nflId == gunner2)
  
  specialistData$topSpeed[i] <- max(gunner1Frames$s)
  specialistData$topSpeed[i + 1] <- max(gunner2Frames$s)
}
```

## 3.13 Fill *squeezeDis* Column
Once the ball is caught, gunners are trained to move towards the center of the field and push the returner towards the sideline. This creates a net for the rest of the team to trap the returner in. This measurement attempts to show how well a gunner is able to push the returner to the sideline by taking the difference between the returners position when they caught the ball, and their average position after they caught the ball.
```{r}
for (i in seq(1, nrow(specialistData), 2)) {
  play <- puntPlays %>% filter(gameId == specialistData$gameId[i] &
                               playId == specialistData$playId[i])
  
  # squeezeDis is only applicable to returns
  if (play$specialTeamsResult[1] == 'Fair Catch') {
    next
  }
  
  returner <- play$returnerId[1]
  
  snapFrame <- puntPlayInfo$snapFrame[which(puntPlayInfo$gameId == play$gameId[1] &
                                              puntPlayInfo$playId == play$playId[1])]
  
  catchFrame <- puntPlayInfo$catchFrame[which(puntPlayInfo$gameId == play$gameId[1] &
                                                puntPlayInfo$playId == play$playId[1])]
  
  # returner pos when ball is caught
  returnerCaughtYPos <- play$y[which(play$nflId == returner & play$frameId == catchFrame)]
  returnerCaughtXPos <- play$x[which(play$nflId == returner & play$frameId == catchFrame)]
  
  # pos of sidelines
  sidelines <- c(0, 53.3)
  
  # determine closest sideline
  dis1 <- abs(returnerCaughtYPos - sidelines[1])
  dis2 <- abs(returnerCaughtYPos - sidelines[2])
  if (dis1 < dis2) {
    sidelineClose <- 0
  } else {
    sidelineClose <- 53.3
  }
  
  # returner pos after ball is caught
  returnerYPos <- play$y[which(play$nflId == play$returnerId & play$frameId > catchFrame)]
  # how far is returner from sideline when they catch ball
  returnerDisSidelineCaught <- abs(returnerCaughtYPos - sidelineClose)
  # mean pos during return
  returnerMeanYPos <- mean(play$y[which(play$nflId == returner & play$frameId > catchFrame)])

  # mean distance to closest sideline during return
  avgSidelineDis <- abs(returnerMeanYPos - sidelineClose)

  specialistData$squeezeDis[c(i, i+1)] <- returnerDisSidelineCaught - avgSidelineDis
}

specialistData$squeezeDis <- ifelse(specialistData$specialTeamsResult == 'Fair Catch', NA, specialistData$squeezeDis)
```

## 3.14 Record if Gunner Made a Tackle
```{r}
tackle <- numeric(nrow(specialistData))
for (i in seq(1, nrow(specialistData), 2)) {
  play <- puntPlays %>% filter(gameId == specialistData$gameId[i] &
                               playId == specialistData$playId[i])
  
  gunner1 <- specialistData$gunnerId[i]
  gunner2 <- specialistData$gunnerId[i + 1]
  tackler <- play$tackler[1]
  
  # If tackle was made, grab player ID of tackler
  if (!(is.na(tackler))) {
    tacklerTeam <- gsub(" .*$", "", tackler)
    tacklerNumber <- substrRight(tackler, 2)
    
    tacklerID <- play$nflId[which(play$teamAbbr == tacklerTeam & 
                                  play$jerseyNumber == tacklerNumber)][1]
    
    # skip play if tackler can't be ID
    if (is.na(tacklerID)) {
      next
    }
    
    # check if tackler ID matches either gunner ID
    if (tacklerID == gunner1) {
      tackle[i] <- 1
    } else if (tacklerID == gunner2) {
      tackle[i + 1] <- 1
    }
  }
}
specialistData$tackle <- tackle
```

## 3.15 Record if Gunner Missed a Tackle
```{r}
missedTackle <- numeric(nrow(specialistData))
for (i in seq(1, nrow(specialistData), 2)) {
  play <- puntPlays %>% filter(gameId == specialistData$gameId[i] &
                               playId == specialistData$playId[i])
  
  gunner1 <- specialistData$gunnerId[i]
  gunner2 <- specialistData$gunnerId[i + 1]
  missedTackler <- play$missedTackler[1]
  
  # If missed tackle occurred, grab player ID of missed tackler
  if (!(is.na(missedTackler))) {
    missedTacklerTeam <- gsub(" .*$", "", missedTackler)
    missedTacklerNumber <- substrRight(missedTackler, 2)
    
    missedTacklerID <- play$nflId[which(play$teamAbbr == missedTacklerTeam & 
                                        play$jerseyNumber == missedTacklerNumber)][1]
    
    # skip play if player can't be ID
    if (is.na(missedTacklerID)) {
      next
    }
    
    # check if ID matches either gunner ID
    if (missedTacklerID == gunner1) {
      missedTackle[i] <- 1
    } else if (missedTacklerID == gunner2) {
      missedTackle[i + 1] <- 1
    }
  }
}
specialistData$missedTackle <- missedTackle
```

## 3.16 Record Gunner Release Types
A release is essentially how the gunner moves off of the line of scrimmage. This can be classified in two ways. One way is either an inside or outside release. An inside release is when the gunner begins moving towards the center of the field, and an outside release is when they start moving towards the sideline. Another way I classified each gunner's release was as correct or incorrect. A correct release would be releasing in the direction that gives the gunner the most direct route to the ball.

### 3.16.1 Record Which Side of the Field Each Gunner is Lined Up.
```{r}
for (i in seq(1, nrow(specialistData), 2)) {
  play <- puntPlays %>% filter(gameId == specialistData$gameId[i] &
                               playId == specialistData$playId[i])
  
  # ID players
  gunner1 <- specialistData$gunnerId[i]
  gunner2 <- specialistData$gunnerId[i + 1]
  
  # get y pos of gunners at start
  gunner1StartYPos <- play$y[which(play$nflId == gunner1 & play$frameId == 1)]
  gunner2StartYPos <- play$y[which(play$nflId == gunner2 & play$frameId == 1)]
  
  # compare y pos to middle of field
  if (gunner1StartYPos > 26.65) {
    specialistData$gunnerSide[i] <- 'L'
  } else {
    specialistData$gunnerSide[i] <- 'R'
  }
  
  if (gunner2StartYPos > 26.65) {
    specialistData$gunnerSide[i + 1] <- 'L'
  } else {
    specialistData$gunnerSide[i + 1] <- 'R'
  }
}
```

### 3.16.2 Classify as Inside or Outside
```{r}
for (i in 1:nrow(specialistData)) {
  play <- puntPlays %>% filter(gameId == specialistData$gameId[i] &
                               playId == specialistData$playId[i])
  
  # ID players
  gunnerID <- specialistData$gunnerId[i]
  viseID <- specialistData$viseId[i]
  
  snapFrame <- puntPlayInfo$snapFrame[which(puntPlayInfo$gameId == play$gameId[1] &
                                              puntPlayInfo$playId == play$playId[1])]
  # gunner y pos at start
  gunnerStartYPos <- play$y[which(play$nflId == gunnerID & play$frameId == snapFrame)]
  
  # gunner and jammer y pos 3 sec after ball is snapped
  # pos taken after 3 sec to make sure gunner isn't just trying to misdirect jammer
  gunnerReleaseYPos <- play$y[which(play$nflId == gunnerID & play$frameId == snapFrame + 30)]
  viseReleaseYPos <- play$y[which(play$nflId == viseID & play$frameId == snapFrame + 30)]
  
  # left side of field , right of jammer
  if (gunnerStartYPos < 26.65 & gunnerReleaseYPos > viseReleaseYPos) {
    specialistData$release[i] <- 1 # inside
  }
  
  # left side of field, left of jammer
  if (gunnerStartYPos < 26.65 & gunnerReleaseYPos < viseReleaseYPos) {
    specialistData$release[i] <- 0 # outside
  }
  
  # right side of field, right of jammer
  if (gunnerStartYPos > 26.65 & gunnerReleaseYPos > viseReleaseYPos) {
    specialistData$release[i] <- 0 # outside
  }
  
  # right side of field, left of jammer
  if (gunnerStartYPos > 26.65 & gunnerReleaseYPos < viseReleaseYPos) {
    specialistData$release[i] <- 1 # inside
  }
}
```

### 3.16.3 Record Kick Direction Relative to Gunner's Position
The kick can end up:

* SS - On the same side of the field as the gunner
* C - In the center of the field
* OS - On the opposite side of the field as the gunner
```{r}
specialistData$kickDirRelGun <- ifelse(specialistData$gunnerSide == specialistData$kickDir, 
                                       'SS', 
                                       specialistData$kickDirRelGun)

specialistData$kickDirRelGun <- ifelse(specialistData$kickDir == 'C', 
                                       'C', 
                                       specialistData$kickDirRelGun)

specialistData$kickDirRelGun <- ifelse(specialistData$kickDir != 'C' & 
                                         specialistData$gunnerSide != specialistData$kickDir,
                                       'OS', 
                                       specialistData$kickDirRelGun)

specialistData$kickDirRelGun[which(is.na(specialistData$kickDir))] <- NA
```

### 3.16.4 Categorize Each Release Type
* 0 - Outside release, kick on same side
* 1 - Outside release, kick on opposite side
* 2 - Outside release, kick to center
* 3 - Inside release, kick to center
* 4 - Inside release, kick on opposite side
* 5 - Inside release, kick on same side
```{r}
specialistData$cat[which(specialistData$release == 0 & specialistData$kickDirRelGun == 'SS')] <- 0
specialistData$cat[which(specialistData$release == 0 & specialistData$kickDirRelGun == 'OS')] <- 1
specialistData$cat[which(specialistData$release == 0 & specialistData$kickDirRelGun == 'C')] <- 2
specialistData$cat[which(specialistData$release == 1 & specialistData$kickDirRelGun == 'C')] <- 3
specialistData$cat[which(specialistData$release == 1 & specialistData$kickDirRelGun == 'OS')] <- 4
specialistData$cat[which(specialistData$release == 1 & specialistData$kickDirRelGun == 'SS')] <- 5
```

### 3.16.5 Categorize Each Release as Correct or Incorrect
If the ball is kicked to the opposite side of the field, the gunner should release inside. If it's on the same side, they should release outside. These routes give them a direct path to the ball.
```{r}
specialistData$correctRelease[which(specialistData$cat == 0 | 
                                      specialistData$cat == 4 | 
                                      specialistData$cat == 3)] <- 1

specialistData$correctRelease[which(specialistData$cat == 1 | 
                                      specialistData$cat == 5 | 
                                      specialistData$cat == 2)] <- 0
```

## 3.17 Remove Unnecessary Variables
```{r}
specialistData <- specialistData[c("gameId", "playId", "gunnerTeam", "gunnerId",
                                   "gunnerName","viseTeam", "viseId", "viseName", 
                                   "timeToBeatVise", "firstManDown", "disFromLOS", 
                                   "disFromReturner", "speedDev", "topSpeed", 
                                   "tackle", "missedTackle", "squeezeDis", "returnYds", 
                                   "specialTeamsResult", "release", "correctRelease")]
```

## 3.18 Write to .csv
```{r}
write.csv(specialistData, file = paste0(wd$data, 'specialist_data.csv'), row.names = FALSE)
```

